---
title: "ARIMA forecasts"
output: html_notebook
---

Using the auto.arima function in the forecast package to select automatically the optimal ARIMA model. Then creates dynamic forecast and additionally static forecast (rolling window) for comparison.


Load data:
```{r}
library(readr)
library(forecast, quietly=TRUE)
library(ggplot2,quietly=True)
library(lubridate)
#library(forecastHybrid,quietly=True)

# Load data
setwd('/Users/david/AnacondaProjects/lstm-load-forecasting')
ds <- read_delim("./data/fulldataset.csv", ";", escape_double = FALSE, col_types = cols(X1 = col_datetime(format = "%Y-%m-%d %H:%M:%S%z")), trim_ws = TRUE)
colnames(ds)[1] <- "time"
```
Split the data into train and test. Cutoff date is 01-02-2017.
```{r}
cutoff <- as.POSIXct("2017-02-01 00:00:00", tz="Europe/Zurich")
attributes(cutoff)$tzone <- "UTC"  
training <- subset(ds, ds$time<cutoff)
test <- subset(ds, ds$time>=cutoff)

training <- subset(training, select = -c(weekday_0, hour_0, month_1))
test <- subset(test, select = -c(weekday_0, hour_0, month_1))
training_ex <- subset(training, select = c(3, 36:76))
test_ex <- subset(test, select = c(3, 36:76))
```

Fit model with automatic model selection
```{r}
fit <- auto.arima(training$actual, xreg=training_ex, seasonal=FALSE)
```

Forecast on test data (dynamic forecast)
```{r}
fc <- forecast(fit, xreg=test_ex)
plot(fc, ylim = c(0,12000))
write.csv(fc$mean, file='./data/arma_fc.csv')
```

Forecast on test data (rolling window forecast)
```{r}
starting_utc <- tail(training$time + hours(1), n=1)
ending_utc <- tail(test$time, n=1)
starting_loc <- starting_utc
ending_loc <- ending_utc
attributes(starting_loc)$tzone <- "Europe/Zurich"
attributes(ending_loc)$tzone <- "Europe/Zurich"

fc_days <- seq(starting_loc, ending_loc, "DSTday")
predictions = seq(starting_loc, ending_loc, "hour")
attributes(predictions)$tzone <- "UTC"

# Store the order terms
order <- arimaorder(fit)
```


```{r}
fc_days_utc <- seq(starting_utc, ending_utc, "DSTday") 
for(i in seq_along(fc_days_utc))
{
  hour_utc <- fc_days_utc[i]
  attributes(hour_utc)$tzone <- "UTC"
  
  # The first iteration can just use the previous estimated model
  if(i == 1){
    l_test <- subset(ds, (ds$time < hour_utc + days(2) & ds$time >= hour_utc))
    l_test_ex <- subset(l_test, select = c(3, 36:76))
    results <- forecast(fit, xreg=l_test_ex)$mean
    pred <- results
    next
  }
  
  # Skip if last day
  if(i > 104){
    next
  }

  # Select window for reestimating model (last 24 hours, train) and forecasting (next 48 hours, test)
  l_training <- subset(ds, ds$time<hour_utc)
  l_test <- subset(ds, (ds$time < hour_utc + days(2) & ds$time >= hour_utc))
  l_training <- subset(l_training, select = -c(weekday_0, hour_0, month_1))
  l_test <- subset(l_test, select = -c(weekday_0, hour_0, month_1))
  l_training_ex <- subset(l_training, select = c(3, 36:76))
  l_test_ex <- subset(l_test, select = c(3, 36:76))
  
  # Reestimate model with new data
  refit <- Arima(l_training$actual, order[1:3], seasonal=FALSE, xreg=l_training_ex)
  results <- forecast(refit, xreg=l_test_ex)$mean
  
  # Actually the t+25:t+48 forecasts are required for comparison with ENTSOE
  pred <- c(pred, results[25:48])
}
write.csv(pred, file="./data/arma_rolling_fc.csv")
```

Now with all data
```{r}
training_ex <- subset(training, select = c(4:11, 12,13,15,16,18,19,21,22,24,25,27,28,30,31,33,34,36:76))
test_ex <- subset(test, select = c(4:11, 12,13,15,16,18,19,21,22,24,25,27,28,30,31,33,34,36:76))

fit2 <- auto.arima(training$actual, xreg=training_ex, seasonal=FALSE)
fc2 <- forecast(fit2, xreg=test_ex)
plot(fc2, ylim = c(0,12000))
write.csv(fc2$mean, file='./data/arma_fc2.csv')

```

```{r}
starting_utc <- tail(training$time + hours(1), n=1)
ending_utc <- tail(test$time, n=1)
starting_loc <- starting_utc
ending_loc <- ending_utc
attributes(starting_loc)$tzone <- "Europe/Zurich"
attributes(ending_loc)$tzone <- "Europe/Zurich"

fc_days <- seq(starting_loc, ending_loc, "DSTday")
predictions = seq(starting_loc, ending_loc, "hour")
attributes(predictions)$tzone <- "UTC"

# Store the order terms
order <- arimaorder(fit2)


```
Loop through windows
```{r}
fc_days_utc <- seq(starting_utc, ending_utc, "DSTday") 
for(i in seq_along(fc_days_utc))
{
  hour_utc <- fc_days_utc[i]
  attributes(hour_utc)$tzone <- "UTC"
  
  # The first iteration can just use the previous estimated model
  if(i == 1){
    l_test <- subset(ds, (ds$time < hour_utc + days(2) & ds$time >= hour_utc))
    l_test_ex <- subset(l_test, select = c(4:11, 12,13,15,16,18,19,21,22,24,25,27,28,30,31,33,34,36:76))
    results <- forecast(fit2, xreg=l_test_ex)$mean
    pred <- results
    next
  }
  
  # Skip if last day
  if(i > 104){
    next
  }

  # Select window for reestimating model (last 24 hours, train) and forecasting (next 48 hours, test)
  l_training <- subset(ds, ds$time<hour_utc)
  l_test <- subset(ds, (ds$time < hour_utc + days(2) & ds$time >= hour_utc))
  l_training <- subset(l_training, select = -c(weekday_0, hour_0, month_1))
  l_test <- subset(l_test, select = -c(weekday_0, hour_0, month_1))
  l_training_ex <- subset(l_training, select = c(4:11, 12,13,15,16,18,19,21,22,24,25,27,28,30,31,33,34,36:76))
  l_test_ex <- subset(l_test, select = c(4:11, 12,13,15,16,18,19,21,22,24,25,27,28,30,31,33,34,36:76))
  
  # Reestimate model with new data
  refit <- Arima(l_training$actual, order[1:3], seasonal=FALSE, xreg=l_training_ex)
  results <- forecast(refit, xreg=l_test_ex)$mean
  
  # Actually the t+25:t+48 forecasts are required for comparison with ENTSOE
  pred2 <- c(pred2, results[25:48])
}
write.csv(pred2, file="arma_rolling_fc2.csv")
```

```

